---
title: "Credit Card Default Factors in Taiwan"
author: "Shira Stites and Marc Al-Helou"
date: "January 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Objective

We plan to use a data set with default data on Taiwanese clients for 6 months in 2005.
We will look at what set of demographic and financial factors are most predictive of default probability. 


```{r cars}
summary(cars)
```

## Process

1) Define the variables and make sure they are metric 
2) Scale variables if necessary
3) Conduct separate analyses of 3 sets of variables (demographic, financial, combined)
4) Look at correlations of variables
5) Conduct a factor analysis including dimensionality reduction
6) Determine relevent factors
7) Derive predictive regression model using each set of variables (demographic, financial, combined)
8) See which model is the most predictive


ProjectData <- read.csv("UCI_Credit_Card.csv")


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
